---
title: "Linear Model Project in R"
subtitle: "M1–MIDO, 2025–2026"
author:
  - name: "Victor Hadida - Antoine Le Gouz de Saint-Seine"
    affiliation: "Université Paris-Dauphine-PSL"
date: last-modified
lang: en
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: body
    toc-title: "Contents"
    number-sections: false
    smooth-scroll: true
    page-layout: full
    linestretch: 1.2
    fontsize: 125%
    self-contained: true
    df-print: default
    code-fold: true 
    code-tools: false
    code-line-numbers: false
    code-overflow: scroll
    code-copy: true
    highlight-style: github
    theme: cosmo
    code-block-bg: "#F5F5F5" # Choose here https://www.w3schools.com/colors/colors_shades.asp
    html-math-method: mathjax
    fig-format: retina
    fig-width: 13
    fig-height: 7
    fig-align: center
    fig-pos: "H"
    fig-cap-location: bottom
    mainfont: "Source Sans Pro" # "Source Sans Pro" "Arial"  "Calibri"  "Cambria"  "Times New Roman"  "Georgia",  your favorite
execute:
  echo: true
  warning: false
  message: false
  error: false
editor:
  markdown:
    wrap: 110
    canonical: true
knitr:
  opts_chunk:
    tidy: styler
    comment: ''
    dev: png
---

```{r}
#| label: setup
#| include: false

# Reproducibility
set.seed(42)

# Global options
options(pillar.width = 130, width = 200, scipen = 999, digits = 5)

# Add Packages here. Be careful of the ordering
library(MASS)
library(jtools)
library(huxtable)
library(patchwork)
library(janitor)
library(broom)
library(performance)
library(parameters)
library(datawizard)
library(see)
library(effectsize)
library(insight)
library(correlation)
library(modelbased)
library(glue)
library(scales)
library(GGally)
library(ggpubr)
library(car)
library(lmtest)
library(rstatix)
library(matrixTests)
library(ggfortify)
library(qqplotr)
library(collapse)
library(kableExtra)
library(tidyverse) # always load it in last position
```

# Introduction

Ce projet étudie les facteurs influençant la note finale de 5000 étudiants via un modèle de régression
linéaire multiple. L'objectif est de comprendre l'impact de variables (numériques) comme les heures d'étude ou
(catégorique) la qualité du sommeil, tout en créant un modèle capable de prédire les scores avec précision.

# Data Management & Exploration

## Data Dictionary & Pre-processing

```{r}
raw_data <- read_csv("project.csv", show_col_types = FALSE)
#Cleaning the raw_data

data_clean <- raw_data|>
  mutate(
    sexe= factor(sexe, levels = 1:3, labels = c("Female", "Male", "Other")),
    
    agecat = factor(agecat, levels = 1:5, labels = c("[14.0, 15.]", "[15.1, 16.1]", "[16.1, 17,1.]", "[17,1, 18.1]","[18,1, 19]") ),
    
    school_type = factor(school_type, levels = 1:2, labels =c("Public", "Private")),
    
    parent_educ = factor(parent_educ, levels = 1:6, labels =c("No Formal", "High School", "Graduate", "Post Graduate 1", "Post Graduate 2", "PHD")),
    
    sleep_qual = factor(sleep_qual, levels= 1:3, labels = c("Poor", "Average", "Good")),
    
    attend_pct_cat = factor(attend_pct_cat, levels = 1:4, labels = c("[50, 62[","[62, 72[","[72, 81[","[81, 100[")),
    
    web_access = factor(web_access, levels = 1:2, labels = c("No", "Yes")),
    
    trav_time = factor(trav_time, levels = 1:4, 
                       labels = c("<15 Min","15-30 Min" ,"30-60 Min", ">60 Min")),
    
    extra_act = factor(extra_act, levels = 1:2, labels = c("No", "Yes")),
    
    study_method = factor(study_method, levels = 1:6, labels = c("Online Videos", "Coaching", "Notes", "Textbook", "Group Study", "Mixed"))
  )

#Relabelling it 
data_clean <- data_clean |>
  relabel(
    id = "ID", y ="Grade", age = "Age", agecat = "Category of age", sexe = "Gender", school_type = "School Type", parent_educ = "Parental Education", 
    study_hrs= "Weekly study Hours ", sleep_hrs = "Sleep Duation (hours) ", sleep_qual = "Sleep Quality", attend_pct = "School attendance (%)", 
    attend_pct_cat = "Categorical School attendance (%)", web_access = "Web access", trav_time = "Commute time", extra_act = "Extracurricular activities",
    study_method = "Study Method"
  )

  
data_clean |>
  
  purrr::map_dfr(~ tibble(
    Type = class(.x)[1],          # Variable type (numeric, factor, etc.)
    N = sum(!is.na(.x)),          # Number of valid observations
    Distinct = n_distinct(.x),    # Number of unique values
    Missing = sum(is.na(.x))      # Number of missing values
  ), .id = "Variable") |>         # The column name becomes the 'Variable' column
  
  
  knitr::kable(
    caption = "<strong style='font-size: 20px;'> Data Dictionary</strong>",
    align = "lcccc", # Left align first col, Center others
    col.names = c("Variable", "Type", "N (Count)", "Distinct Values", "Missing Values")
  )
```

::::::: grid
:::: g-col-6
::: {.callout-tip appearance="simple"}
## Observations

**`r nrow(data_clean)`** students included in the study.
:::
::::

:::: g-col-6
::: {.callout-tip appearance="simple"}
## Variables

**`r ncol(data_clean)`** features retained for analysis.
:::
::::
:::::::

**Variable details:** The dataset consists of `r paste(unique(sapply(data_clean, class)), collapse = " and ")`
variables, including: *`r paste(names(data_clean), collapse = ", ")`*.

```{r}
data_clean |>
  select(where(is.factor)) |>
  map_dfr(~ tibble(Niveaux = paste(levels(.x), collapse = ", ")), .id = "Variable") |>
  
  knitr::kable(caption = "<strong style='font-size: 20px;'>describing factor levels</strong>")
```

::: {.callout-note title="Choice of Reference Levels"}
To ensure interpretability, we have defined the **reference level** (baseline) for each categorical variable.

-   **Strategy:** For ordinal variables (e.g., parent education, age groups), the **lowest rank** was
    systematically selected as the reference. For non ordinal variables we chose the level with the highest
    number of occurence as reference.
:::

```{r}
data_clean <- data_clean |>
  mutate(
    # Only applying frequency reordering to Nominal variables
    sexe = fct_infreq(sexe),
    school_type = fct_infreq(school_type),
    web_access = fct_infreq(web_access),
    extra_act = fct_infreq(extra_act),
    study_method = fct_infreq(study_method)
    # Note: agecat, parent_educ, etc. are NOT touched here to preserve order
  )
#Checking the reference Categories 
data_clean |>
  select(where(is.factor)) |>
  map_dfr(~ tibble(Reference_Category = levels(.x)[1]), .id = "Variable") |>
  knitr::kable(caption = "<strong style='font-size: 20px;'>Reference level of each factor</strong>")
```

## Data Integrity Check

To ensure the reliability of our analysis, we performed a strict data integrity check. We sequentially
addressed duplicate entries and missing values. The cleaning process is summarized below:

```{r}
# 1. Initial count
n_init <- nrow(data_clean)

# 2. Handling Duplicates
data_clean <- data_clean |> distinct()
n_after_dedup <- nrow(data_clean)
n_duplicates <- n_init - n_after_dedup

# 3. Handling Missing Values (NA)
data_clean <- data_clean |> drop_na()
n_final <- nrow(data_clean)
n_na_removed <- n_after_dedup - n_final

# Total removed
n_total_removed <- n_init - n_final
pct_removed <- round((n_total_removed / n_init) * 100, 1)
```

::: {.callout-warning icon="true"}
## Data Cleaning Report

We processed the initial dataset of **`r n_init`** observations:

1.  **Duplicate Detection:** `r n_duplicates` duplicate rows were identified and removed.
2.  **Missing Values:** `r n_na_removed` rows containing missing values (NA) were excluded.

**Impact:** A total of **`r n_total_removed`** observations were removed (`r pct_removed`% of the original
data). The final clean dataset contains **`r n_final`** observations.
:::

## Outlier Detection

Now we focus on detecting potential outliers and extreme values. We use boxplots for continuous variables and
the `summarise` function to verify the coherence of the data ranges (Min-Max check).

```{r}
# 1. Coherence Check (Min-Max)
data_clean |>
  select(where(is.numeric)) |>
  pivot_longer(everything(), names_to = "Variable", values_to = "Valeur") |>
  group_by(Variable) |>
  summarise(
    Minimum = min(Valeur, na.rm = TRUE),
    Maximum = max(Valeur, na.rm = TRUE)
  ) |>
  # Simple kable sans fioritures
  knitr::kable(
    caption = "<strong style='font-size: 20px;'>Range Verification (Min-Max)</strong>",
    digits = 2,
    align = "lcc" 
  )

# 2. Visual Inspection (Boxplots)
vars_num <- c("y", "age", "study_hrs", "sleep_hrs", "attend_pct")

data_clean |> 
  select(all_of(vars_num)) |> 
  pivot_longer(everything(), names_to = "var", values_to = "value") |> 
  ggplot(aes(x = var, y = value)) +
  facet_wrap(~ var, scales = "free", ncol = 3) + 
  
  geom_boxplot(fill = "orange", outlier.colour = "red", outlier.size = 2, alpha = 0.7) + 
  
  theme_bw(base_size = 12) +
  labs(title = "Outliers Detection", x = "", y = "Values") +
  
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(size = 20, face = "bold", hjust = 0.5))

# 3. List of Identified Outliers
# Raw listing of observations flagged as outliers
outliers_table <- data_clean |>
  select(id, all_of(vars_num)) |>
  pivot_longer(-id, names_to = "variable", values_to = "value") |>
  group_by(variable) |>
  rstatix::identify_outliers(value)

n_outliers <- nrow(outliers_table)
n_extreme <- sum(outliers_table$is.extreme)

outliers_details <- outliers_table |>
  count(variable) |>
  mutate(text = paste0("**", variable, "** (", n, ")")) |>
  pull(text) |>
  paste(collapse = ", ")
```

::: {.callout-warning icon="true"}
## Outliers Cleaning Report

We processed the initial dataset:

1.  **Outliers Detection:** `r n_outliers` outliers where detected. They are distributed as follows:
    `r outliers_details`.
2.  **Outlier out of acceptance range** `r n_extreme` occurence with extreme values.
:::

## Feature Selection & Redundancy Analysis

We identified potential redundancies in the dataset where similar information is represented in both
continuous and categorical forms. Specifically, we examined three pairs:

Age: age (continuous) vs agecat (categorical).

Attendance: attend_pct (continuous) vs attend_pct_cat (categorical).

Sleep: sleep_hrs (continuous) vs sleep_qual (categorical).

To determine which form to retain, we analyzed the relationships between these pairs using boxplots. Our goal
was to assess whether the categorical variables provide distinct information from their continuous
counterpart.

```{r}

p1 <- ggplot(data_clean, aes(x = agecat, y = age, fill = agecat)) +
  geom_boxplot(alpha = 0.6, outlier.colour = "red", size = 0.3) +
  scale_fill_brewer(palette = "Blues") +
  labs(title = "agecat vs age", 
       y = "Age (Continuous)", 
       x = "") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none", 
        axis.text.x = element_text(angle = 30, hjust = 1))


p2 <- ggplot(data_clean, aes(x = sleep_qual, y = sleep_hrs, fill = sleep_qual)) +
  geom_boxplot(alpha = 0.6, outlier.colour = "red", size = 0.3) +
  scale_fill_brewer(palette = "Oranges") +
  labs(title = "sleep_qual vs sleep_hrs", 
       y = "sleep_hrs ", 
       x = "") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 30, hjust = 1))


p3 <- ggplot(data_clean, aes(x = attend_pct_cat, y = attend_pct, fill = attend_pct_cat)) +
  geom_boxplot(alpha = 0.6, outlier.colour = "red", size = 0.3) +
  scale_fill_brewer(palette = "Greens") +
  labs(title = "attend_pct_cat vs attend_pct", 
       y = "attend_pct (Continuous)", 
       x = "") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 30, hjust = 1))


figure_finale <- ggarrange(p1, p2, p3, ncol = 3, nrow = 1)

annotate_figure(figure_finale,
                top = text_grob("Redundancy Analysis: Coutinuous vas Categorical ", 
                                face = "bold", size = 14))
```

::: {.callout-important icon="true"}
## Decision: Data Management Strategy

Following a visual inspection of associations using boxplots, we proceeded with the following simplifications:

Retention of sleep_qual: The boxplots revealed considerable overlap between sleep duration and quality
ratings, suggesting that sleep_qual captures subjective information distinct from sleep_hrs.

Removal of Redundant Factors: Conversely, age and attend_pct showed perfect separation from their categorical
versions (agecat, attend_pct_cat) with no overlap. To avoid redundancy and multicollinearity, we removed these
categorical variables in favor of their continuous counterparts. Also the continuous variables yields to
simpler models and capture more information.
:::

```{r}
data_final <- data_clean |>
  select(-agecat, -attend_pct_cat)
  
data_final |>
  
  purrr::map_dfr(~ tibble(
    Type = class(.x)[1],          # Variable type (numeric, factor, etc.)
    N = sum(!is.na(.x)),          # Number of valid observations
    Distinct = n_distinct(.x),    # Number of unique values
    Missing = sum(is.na(.x))      # Number of missing values
  ), .id = "Variable") |>         # The column name becomes the 'Variable' column
  
  
  knitr::kable(
    caption = "<strong style='font-size: 20px;'>Final Data Dictionary</strong>",
    align = "lcccc", # Left align first col, Center others
    col.names = c("Variable", "Type", "N (Count)", "Distinct Values", "Missing Values")
  )
```

# Exploratory Data Analysis (EDA)

Before starting this section, we split the data into a training set and a test set using the standard 80-20
partition.

```{r}
idx <- sample(seq_len(nrow(data_final)), size = floor(0.8*nrow(data_final)))
data_train <- data_final[idx, ]
data_test  <- data_final[-idx, ]
```

## Descriptive Statistics

Here is a table describing the major statistics of our quantitative variables.

```{r}
select(data_train, y, age, study_hrs, sleep_hrs, attend_pct) |>
  get_summary_stats(show = c("n", "mean", "sd", "median", "q1", "q3", "min", "max", "iqr"))|>
  knitr::kable(
    caption = "<strong style='font-size: 20px;'>Statistics</strong>",
    align = "llcc")|>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE,
    position = "center"
  )
```

Here are the frequency tables for our categorical variables.

```{r}
vars_cat <- c("sexe", "school_type", "parent_educ", "web_access", 
              "trav_time", "extra_act", "study_method", "sleep_qual")

make_summary <- function(var_name){
  data_train |>
    tabyl(!!sym(var_name)) |>              
    adorn_pct_formatting(digits = 2) |>    
    arrange(desc(n)) |>                    
    rename(Category = 1) |>                
    mutate(Variable = var_name) |>        
    select(Variable, Category, n, percent)
}

map_dfr(vars_cat, make_summary) |>   
  kable(
    caption = "<strong style='font-size: 20px;'>Distribution of Categorical Variables</strong>",
    align = "llcc"
  ) |>
  # Improve Visual: Group rows and add styling
  collapse_rows(columns = 1, valign = "top") |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE, 
    position = "center"
  )
```

## Visual exploration

```{r}
# 1. Calculate Distribution Metrics
s_y <- skewness(data_train$y, na.rm = TRUE)
k_y <- kurtosis(data_train$y, na.rm = TRUE)

# 2. Histogram
plot_hist <- ggplot(data_train, aes(x = y)) + 
  geom_histogram(bins = 15, color = "black", fill = "firebrick", alpha = 0.8) +
  theme_bw(base_size = 12) +
  labs(
    title = "Distribution of Target Variable (y)",
    subtitle = paste0("Skewness: ", round(s_y, 2), " | Kurtosis: ", round(k_y, 2)),
    x = "Final Grade (y)", 
    y = "Frequency"
  )

# 3. Boxplot
plot_box <- ggplot(data_train, aes(x = y)) + 
  # Using x=y creates a horizontal boxplot, which aligns nicely with the histogram's axis
  geom_boxplot(fill = "orange", outlier.colour = "red", outlier.size = 2, alpha = 0.6) + 
  theme_bw(base_size = 12) +
  labs(
    title = "Boxplot of Grades", 
    x = "Final Grade (y)", 
    y = ""
  ) +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())

# 4. Combine Visuals
ggarrange(plot_hist, plot_box, ncol = 2, nrow = 1)
```

::: {.callout-note title="Interpretation"}
First, visual inspection of the boxplot reveals a few outliers. These likely represent edge cases,
corresponding to exceptionally high- or low-performing students. However, the histogram shows that the overall
distribution is fairly symmetric.
:::

## Testing y against Quantitative Predictors

```{r}
data_train |>
  select(y, age, study_hrs, sleep_hrs, attend_pct) |>
  pivot_longer(cols = c(age, study_hrs, sleep_hrs, attend_pct)) |>
  ggplot(aes(x = value, y = y)) +
  # Using 2 columns is better for document layout than a single vertical column
  facet_wrap(vars(name), ncol = 2, scales = "free_x") +
  # Reduced alpha to 0.6 to better visualize overlapping points
  geom_point(size = 2, shape = 21, fill = "red", color = "black", alpha = 0.6) +
  geom_smooth(method = "lm", color = "blue", se = TRUE) +
  labs(x = "Predictor Value", y = "Grade (y)") +
  theme_bw(base_size = 12)
```

::: {.callout-note title="Key Observations"}
Linearity: Visual inspection suggests a general linear relationship between these predictors and the final
grade, supporting the use of linear regression.

Variance (Homoscedasticity): Regarding variance, we observe different behaviors across predictors:

Age and Attendance: The variance appears relatively constant (homoscedastic).

Sleep and Study Hours: We observe very slight signs of heteroscedasticity (a "megaphone" shape), where the
spread of grades increases as the predictor value increases. This behavior will be further monitored in the
residual analysis.

Additionally, the regression line for age is nearly horizontal, suggesting that age won't bring any
information in our regression.
:::

## Testing y against Categorical Predictors

```{r}
data_train |>
  select(y, sexe, school_type, parent_educ, web_access, sleep_qual, trav_time, extra_act, study_method) |>
  pivot_longer(cols = -y) |> # Pivot everything except 'y'
  ggplot(aes(x = value, y = y)) +
  facet_wrap(vars(name), ncol = 4, scales = "free_x") +
  geom_boxplot(fill = "orange", outlier.colour = "red", outlier.size = 2, alpha = 0.6) +
  labs(y = "Grade (y)", x = "Category") +
  theme_bw(base_size = 12) +
  # Rotate labels slightly to prevent overlapping
  theme(axis.text.x = element_text(angle = 15, hjust = 1))
  
```

::: {.callout-note title="Key Observations"}
Sexe: The boxplots for both genders are nearly identical, with medians on the same level and significant
overlap in distribution. This suggests that gender is likely not a useful predictor for determining the grade.

Other Factors: Conversely, factors like parent_educ and study_method show distinct differences in median
grades between levels, indicating potential predictive power.
:::

## Summarizing Assocation

### Associating Numeric with Numeric (Correlation Analysis)

```{r}
 # 1. Visual Correlation Matrix
data_train |>
  select(age, sleep_hrs, study_hrs, attend_pct) |>
  correlation(method = "pearson") |>
  summary(redundant = TRUE) |>
  plot() +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom") +
  labs(title = "Correlation Matrix Heatmap")

# 2. Detailed Table
data_train |>
  select(age, sleep_hrs, study_hrs, attend_pct) |>
  correlation(method = "pearson") |>
  # Convert to standard data frame for cleaner printing
  as.data.frame() |> 
  select(Parameter1, Parameter2, r, p, CI_low, CI_high) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  kbl(caption = "Pearson Correlation Coefficients") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

::: {.callout-note title="Key Insights"}
We examine these correlations to ensure no two predictors are highly correlated (e.g., r=\>0.8), which would
destabilize the model. There are no signs of severe multicollinearity. We can safely include these variables
together in the multiple regression model without fear of redundancy.
:::

### Associating Categorical with Categorical (Interaction Screening)

In order to detect potential valuable associations between two categorical variables, we compared the AIC of
two models: one additive model (e.g. y\~ sexe + sleep_qual) and one with interaction (e.g. y\~ sexe \*
sleep_qual).

```{r}
# 1. Define Screening Function
test_pair <- function(vars, target, data){
  combinaisons <- combn(vars, 2, simplify = FALSE)
  results <- data.frame()
  
  for(pair in combinaisons) {
    v1 <- pair[1]
    v2 <- pair[2]
    
    # Define Formulas
    f_add <- as.formula(paste(target, "~", v1, "+", v2))
    f_int <- as.formula(paste(target, "~", v1, "*", v2))
    
    # Fit Models
    m_add <- lm(f_add, data = data)
    m_int <- lm(f_int, data = data)
    
    # Calculate Delta AIC (Positive = Interaction is better)
    # Delta = AIC(Simple) - AIC(Complex)
    delta <- AIC(m_add) - AIC(m_int) 
    
    results <- rbind(results, data.frame(
      Pair = paste(v1, "*", v2),
      AIC_Additive = AIC(m_add),
      AIC_Interaction = AIC(m_int),
      Delta_AIC = delta))
  }
  
  # Return sorted results
  results |> arrange(desc(Delta_AIC))
}

# 2. Run Screening
cat_vars <- c("sexe", "parent_educ", "school_type", "sleep_qual", 
              "web_access", "trav_time", "extra_act", "study_method")

cat_interactions <- test_pair(cat_vars, "y", data_train)

# 3. Render Professional Table
cat_interactions |>
  # Round for readability
  mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  
  # Conditional Formatting: Green if useful (Delta > 2), Gray otherwise
  mutate(
    Delta_AIC = cell_spec(Delta_AIC, 
                          color = ifelse(Delta_AIC > 2, "green", "gray"),
                          bold = ifelse(Delta_AIC > 2, TRUE, FALSE))
  ) |>
  
  # Create Table
  kable(
    format = "html", 
    escape = FALSE,  # Required for color formatting
    col.names = c("Tested Pair", "AIC (Additive)", "AIC (Interaction)", "Delta AIC"),
    align = "c", 
    caption = "<strong style='font-size: 20px;'>Model Comparison: Additive vs. Interaction</strong>"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE, 
    position = "center"
  ) |>
  scroll_box(height = "400px")
```

::: {.callout-note title="Interpretation: Interaction Screening"}
Conclusion: We observe that for all tested pairs, the Delta AIC is either negative or negligible (below the
threshold of 2).

This indicates that adding interaction terms between categorical variables increases model complexity without
providing a statistically significant improvement in fit. Consequently, the simpler additive model is
preferred in all cases, and we will not include categorical-categorical interactions in our final candidate
models.
:::

### Associating Categorical with Numerical (Mixed Interactions)

We visually screen for interactions by looking for "crossing lines" or significantly different slopes in the
scatterplots. We checked all differents combinations of numerical and categorical variables. Here is a display
of the ones with significant interactions in our eyes.

```{r}
# Plot 1: Sleep Hours vs Grade, colored by Sleep Quality
s1 <- ggplot(data_train, aes(x = sleep_hrs, y = y, color = sleep_qual)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Sleep Hours", y = "Grade", color = "Sleep Qual") +
  theme_bw(base_size = 11) + theme(legend.position = "bottom")

# Plot 2: Sleep Hours vs Grade, colored by Sex
s2 <- ggplot(data_train, aes(x = sleep_hrs, y = y, color = sexe)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Sleep Hours", y = "Grade", color = "Sex") +
  theme_bw(base_size = 11) + theme(legend.position = "bottom")

# Plot 3: Age vs Grade, colored by Study Method
s3 <- ggplot(data_train, aes(x = age, y = y, color = study_method)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Age (Years)", y = "Grade", color = "Method") +
  theme_bw(base_size = 11) + theme(legend.position = "bottom")

# Plot 4: Study Hours vs Grade, colored by Sleep Quality
s4 <- ggplot(data_train, aes(x = study_hrs, y = y, color = sleep_qual)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Study Hours", y = "Grade", color = "Sleep Qual") +
  theme_bw(base_size = 11) + theme(legend.position = "bottom")


ggarrange(s1, s2, s3, s4, ncol = 2, nrow = 2)
```

# Building regression models

## Building Intuition

In order to build on some intuition we modeled some simple models, here is a sample of the summary of three of
them :

```{r}

mod_age <- lm( y ~ age, data=data_train)


mod_study_hrs <-  lm(y ~ study_hrs, data=data_train)


mod_extra_act <-  lm(y ~ extra_act, data=data_train)


export_summs(mod_age, mod_study_hrs, mod_extra_act,
             model.names = c("Age Only", "Study Hrs Only", "Extra Act Only"),
             statistics = c("N" = "nobs", "R²" = "r.squared"),
             digits = 3)
```

## Building First Models (Variable Selection)

After defining the null model `mod_null` and the full model `mod_full`, we proceed with automatic variable
selection to identify the most parsimonious model. We tested three strategies: **Forward Selection**,
**Backward Elimination**, and **Stepwise (Both)**. Since all three methods converged on the exact same set of
predictors, we display the results for the Forward Selection model below.

```{r}
# 1. Define Baseline Models
mod_null <- lm(y ~ 1, data = data_train)
mod_full <- lm(y ~ ., data = data_train)

# 2. Run Forward Selection
# Note: We use trace=FALSE to prevent the report from filling with log output.
mod_forw <- stepAIC(mod_null, 
    y ~ age + study_hrs + sleep_hrs + attend_pct + sexe + school_type +
        parent_educ + sleep_qual + web_access + trav_time + extra_act + study_method, 
    data = data_train, 
    direction = "forward", 
    trace = FALSE)

# 3. Compare the Full Model vs. the Selected (Forward) Model
summary(mod_forw )
```

::: {.callout-note title="Key Findings"}
Confirmation of EDA: The selected model `mod_forw` automatically excluded age and sexe. This mathematically
confirms our earlier visual inspection (flat regression lines) that these variables do not contribute
predictive power.
:::

## Adding Interactions & Final Evaluation

Based on our EDA, we hypothesized potential interactions (e.g., Sleep Quality moderating Sleep Hours). We now
test adding these interaction terms to our model (`mod_forw`) to see if they improve predictive performance.

### 1. Defining Interaction Models

We proceed in two steps, first creating `mod_forw_inter_1` by adding the four interactions we detected during
the EDA to `mod_forw`. Then after careful examination of the summary of `mod_forw_inter_1` (p-values of
students tests), we defined `mod_forw_inter_2` where we only kept intercations **sleep_qual:sleep_hrs** and
**study_hrs:sleep_qual**.

```{r}

# Model 1: Add all potential interactions identified in EDA
mod_forw_inter_1 <- lm(y ~ study_hrs + sleep_hrs + attend_pct + school_type +
    parent_educ + sleep_qual + web_access + trav_time + extra_act + study_method + 
    sleep_qual:sleep_hrs + age:study_method + sexe:sleep_hrs + study_hrs:sleep_qual, 
    data = data_train)

# Model 2: Keep only the statistically significant/logical interactions
mod_forw_inter_2 <- lm(y ~ study_hrs + sleep_hrs + attend_pct + school_type +
    parent_educ + sleep_qual + web_access + trav_time + extra_act + study_method + 
    sleep_qual:sleep_hrs + study_hrs:sleep_qual, 
    data = data_train)


```

#### 2. Performance Evaluation (Train vs. Test)

To objectively select the best model, we compare them based on RMSE (Root Mean Squared Error) on the unseen
Test Set.

```{r}
# Function to calculate metrics
get_model_metrics <- function(model, model_name, data_train, data_test) {
  
  # Predictions
  pred_train <- predict(model, newdata = data_train)
  pred_test  <- predict(model, newdata = data_test)
  
  # Metrics (RMSE)
  rmse_train <- sqrt(mean((data_train$y - pred_train)^2))
  rmse_test  <- sqrt(mean((data_test$y - pred_test)^2))
  
  # Model Properties
  aic_val <- AIC(model)
  r2_adj  <- summary(model)$adj.r.squared
  
  return(data.frame(
    Model            = model_name,
    AIC              = aic_val,
    R2_Adj_Train     = r2_adj,
    RMSE_Train       = rmse_train,
    RMSE_Test        = rmse_test
  ))
}

# Compile results
performance_table <- rbind(
  get_model_metrics(mod_full, "mod_full", data_train, data_test),
  get_model_metrics(mod_null, "mod_null", data_train, data_test),
  get_model_metrics(mod_forw, "mod_forw", data_train, data_test),
  get_model_metrics(mod_forw_inter_1, "Interactions (Complex)", data_train, data_test),
  get_model_metrics(mod_forw_inter_2, "Interactions (Optimized)", data_train, data_test)
)

# Render the Comparison Table
performance_table |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |> 
  kable(
    caption = "<strong style='font-size: 20px;'Performance Comparison: Training vs. Test Set</strong>",
    col.names = c("Model", "AIC", "Adj R² (Train)", "RMSE (Train)", "RMSE (Test)"),
    align = "lcccc"
  ) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
  column_spec(5, bold = TRUE, background = "#e6f2ff", border_left = TRUE)


# ANOVA Test: Does the complex Model 1 offer a significant improvement over Model 2?
# Result: If p > 0.05, we prefer the simpler model (Model 2).
anova_results <- anova(mod_forw_inter_1, mod_forw_inter_2)

# 3. Rename rows to match the order above
# Row 1 = First argument (mod_forw_inter_1 -> Complex)
# Row 2 = Second argument (mod_forw_inter_2 -> Optimized)
row.names(anova_results) <- c("1. mod_forw_inter_1", "2. mod_forw_inter_2")

# 4. Display Clean Table
anova_results |>
  kable(
    caption = "<strong style='font-size: 20px;'>ANOVA Test: Model Comparison</strong>",
    digits = 3
  ) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
# Assign the winner
mod_final <- mod_forw_inter_2

```

::: {.callout-note title="Final Model Selection"}
Looking at the results, `mod_forw_inter_1` and `mod_forw_inter_2` demonstrate the best performance criteria.
This confirms the conclusion from our EDA: the model fits better when including the selected interaction
terms.

Since the predictive performance of both models is nearly identical, we prefer `mod_forw_inter_2`. It is a
simpler model and is therefore less prone to overfitting.

To validate this choice, we performed a nested F-test. With a p-value of 0.561 (≫0.05), we fail to reject the
null hypothesis (H0). Consequently, we adopt `mod_forw_inter_2` as our final model, designated as `mod_final`.
:::

# Diagnostics, assumptions, and robustness

## Running some diagnostics

### Residuals vs Fitted

```{r}
autoplot(mod_final, which = 1, ncol = 1, colour = "grey", shape = 21, size = 2) +
  theme_bw(base_size = 14) +
  labs_pubr()
```

::: {.callout-note title="Plot analysis "}
We observe no apparent pattern in the plot, and the trend line is flat near zero. The residuals are centered.
:::

### Residuals vs Predictors (Numerical ones)

```{r}
res_df <- augment(mod_final) %>%
  mutate(resid = .resid, fitted = .fitted)

p_res1 <- ggplot(res_df, aes(x = study_hrs, y = resid)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(shape = 21, size = 2, colour = "grey") +
  geom_smooth(method = "loess", se = FALSE) +
  labs(x = "Study hours", y = "Residuals", title = "Residuals vs Study hours") +
  theme_bw(base_size = 14) +
  ggpubr::theme_pubr()

p_res2 <- ggplot(res_df, aes(x = sleep_hrs, y = resid)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(shape = 21, size = 2, colour = "grey") +
  geom_smooth(method = "loess", se = FALSE) +
  labs(x = "Study hours", y = "Residuals", title = "Residuals vs Study hours") +
  theme_bw(base_size = 14) +
  ggpubr::theme_pubr()

p_res3 <- ggplot(res_df, aes(x = attend_pct, y = resid)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(shape = 21, size = 2, colour = "grey") +
  geom_smooth(method = "loess", se = FALSE) +
  labs(x = "Study hours", y = "Residuals", title = "Residuals vs Study hours") +
  theme_bw(base_size = 14) +
  ggpubr::theme_pubr()

p_res1 + p_res2 + p_res3
```

::: {.callout-note title="Plot analysis "}
We observe no apparent pattern in the plots, and the trend line is flat near zero. The residuals are centered.
:::

### Residuals vs Predictors (Categorical ones)

```{r}
# Define a consistent theme for all plots
my_theme <- theme_pubr() + 
            theme(legend.position = "none") # Hide legend (redundant with x-axis)

# 1. School Type
pres_4 <- ggplot(res_df, aes(x = school_type, y = resid, fill = school_type)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_boxplot(alpha = 0.6) + # Alpha makes colors softer
  labs(x = "School Type", y = "Residuals", title = "vs. School Type") +
  my_theme

# 2. Parent Education
pres_5 <- ggplot(res_df, aes(x = parent_educ, y = resid, fill = parent_educ)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_boxplot(alpha = 0.6) +
  labs(x = "Parent Education", y = "Residuals", title = "vs. Parent Education") +
  my_theme

# 3. Sleep Quality
pres_6 <- ggplot(res_df, aes(x = sleep_qual, y = resid, fill = sleep_qual)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_boxplot(alpha = 0.6) +
  labs(x = "Sleep Quality", y = "Residuals", title = "vs. Sleep Quality") +
  my_theme

# 4. Web Access
pres_7 <- ggplot(res_df, aes(x = web_access, y = resid, fill = web_access)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_boxplot(alpha = 0.6) +
  labs(x = "Web Access", y = "Residuals", title = "vs. Web Access") +
  my_theme

# 5. Travel Time
pres_8 <- ggplot(res_df, aes(x = trav_time, y = resid, fill = trav_time)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_boxplot(alpha = 0.6) +
  labs(x = "Travel Time", y = "Residuals", title = "vs. Travel Time") +
  my_theme

# 6. Extra Activities
pres_9 <- ggplot(res_df, aes(x = extra_act, y = resid, fill = extra_act)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_boxplot(alpha = 0.6) +
  labs(x = "Extra Activities", y = "Residuals", title = "vs. Extra Activities") +
  my_theme

# 7. Study Method
pres_10 <- ggplot(res_df, aes(x = study_method, y = resid, fill = study_method)) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_boxplot(alpha = 0.6) +
  labs(x = "Study Method", y = "Residuals", title = "vs. Study Method") +
  my_theme

# Combine plots (2 columns is usually easier to read than one long column)
(pres_4 + pres_5) / (pres_6 + pres_7) / (pres_8 + pres_9) / pres_10
```

::: {.callout-note title="Box-Plot analysis "}
Observation: Across all categorical predictors, the boxplots show that the residuals are centered around zero
. The medians align closely with the zero line, and no single category shows a drastic shift up or down.
:::

### Residuals vs Ommitted variables (Age + Sexe)

```{r}
res_df2 <- data_train |>
  mutate(resid = resid(mod_final))

# 2. Plot 1: Age (Quantitative)
p11 <- ggplot(res_df2, aes(x = age, y = resid)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(shape = 21, size = 2, colour = "grey", fill = "grey", alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(x = "Age (Omitted)", y = "Residuals", title = "vs. Age") +
  theme_pubr()

# 3. Plot 2: Sex (Categorical) -> Styled to match previous plots
p12 <- ggplot(res_df2, aes(x = sexe, y = resid, fill = sexe)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(alpha = 0.6) +
  labs(x = "Sex (Omitted)", y = "Residuals", title = "vs. Sex") +
  theme_pubr() +
  theme(legend.position = "none") # Hide legend as X-axis is sufficient

# 4. Combine
p11 + p12
```

::: {.callout-note title="Plot analysis & Box-Plot analysis "}
Visual inspection of the residuals against the excluded variables (age and sex) reveals no discernible
patterns or bias. This confirms that these factors hold no additional predictive value and validates our
decision to exclude them from the final model.
:::

### Homoskedastic reisudals: Scale-Location Plot + Breusch-Pagan

```{r}
autoplot(mod_final, which = 3, ncol = 1, colour = "grey", shape = 21, size = 2) +
  theme_bw(base_size = 14) +
  labs_pubr()

bptest(mod_final, studentize = FALSE)
```

::: {.callout-note title="Plot analysis & Box-Plot analysis "}
Inspecting the Scale-Location plot, we observe no discernible pattern, and the smoothing line remains
approximately horizontal. This indicates that the assumption of homoscedasticity is satisfied. Breusch-Pagan
test confirms it.
:::

### Uncorrelated Residuals: Durbin-Watson

```{r}
dw_result <- durbinWatsonTest(mod_final)

# 2. Extract key values into a data frame
dw_table <- data.frame(
  "Metric" = c("Autocorrelation", "D-W Statistic", "P-Value"),
  "Value"  = c(
    round(dw_result$r, 4),      # The autocorrelation value
    round(dw_result$dw, 4),     # The Durbin-Watson statistic
    signif(dw_result$p, 4)      # The p-value
  )
)

# 3. Create a nice visual table
dw_table |>
  kable(
    caption = "<strong style='font-size: 20px;'>Durbin-Watson Test </strong>",
    align = "lr" # Left align text, Right align numbers
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE,
    position = "center"
  )
```

::: {.callout-note title="Interpretation: Uncorrelation of Residuals"}
Result: The test yields a p-value significantly greater than 0.05 (α=5%).

Durbin-Watson test confirms that the residuals are uncorrelated.
:::

### Normality of Residuals: QQ-plot + Shapiro

```{r}
p_qq <- ggqqplot(residuals(mod_final), 
         title = "Normal Q-Q Plot of Residuals",
         color = "grey30",
         ggtheme = theme_pubr()) 

p_qq

# 2. Statistical Test: Shapiro-Wilk
# We use tidy() to convert the raw test output into a clean dataframe
shapiro_res <- shapiro.test(residuals(mod_final)) |> 
  tidy()

shapiro_res |>
  kable(
    caption = "<strong style='font-size: 20px;'>Shapiro-Wilk Test for Normality</strong>", 
    digits = 4
  ) |>
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

::: {.callout-note title="Interpretation: Normality of Residuals"}
"The Normal Q-Q plot shows residuals falling closely along the reference line within the confidence bands,
while the Shapiro-Wilk test yields a p-value of 0.2155 (p\>0.05). Consequently, we fail to reject the null
hypothesis and conclude that the residuals are normally distributed, satisfying the assumption for linear
regression."
:::

### Outlier/High Leverage

```{r}

# 1. Prepare Data
# augment adds .cooksd, .hat, and .std.resid to the data
inf_data <- augment(mod_final, data = data_train) |>
  mutate(obs_index = row_number())

# 2. Define Thresholds for visual reference
n <- nrow(data_train)
p <- length(coef(mod_final))
cooks_threshold <- 4 / n          # Common rule of thumb for Cook's D
hat_threshold   <- 2 * p / n      # Common rule of thumb for Leverage

# 3. Create Cook's Distance Plot (Influence)
p_cook <- ggplot(inf_data, aes(x = obs_index, y = .cooksd)) +
  geom_bar(stat = "identity", width = 0.5, fill = "steelblue", alpha = 0.6) +
  geom_hline(yintercept = cooks_threshold, linetype = "dashed", color = "red") +
  annotate("text", x = 10, y = cooks_threshold, label = " Threshold (4/n)", 
           vjust = -0.5, color = "red", size = 3) +
  labs(x = "Observation Index", y = "Cook's Distance", title = "Influence (Cook's D)") +
  theme_pubr()

# 4. Create Hat Values Plot (Leverage)
p_hat <- ggplot(inf_data, aes(x = obs_index, y = .hat)) +
  geom_point(color = "darkgreen", alpha = 0.6) +
  geom_segment(aes(x = obs_index, xend = obs_index, y = 0, yend = .hat), color = "darkgreen", alpha = 0.2) +
  geom_hline(yintercept = hat_threshold, linetype = "dashed", color = "red") +
  annotate("text", x = 10, y = hat_threshold, label = " Threshold (2p/n)", 
           vjust = -0.5, color = "red", size = 3) +
  labs(x = "Observation Index", y = "Hat Values (Leverage)", title = "Leverage (Hat Values)") +
  theme_pubr()

# 5. Combine Plots
p_cook + p_hat

outlier_result <- outlierTest(mod_final)
outlier_df <- data.frame(
  "Observation_ID" = names(outlier_result$rstudent),
  "Studentized_Resid" = round(outlier_result$rstudent, 3),
  "Unadjusted_P_Value" = format.pval(outlier_result$p, digits = 3),
  "Bonferroni_P_Value" = format.pval(outlier_result$bonf.p, digits = 3)
)


outlier_df |>
  kable(
    caption = "Bonferroni Outlier Test Results",
    align = "c"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE,
    position = "center"
  ) |>
  row_spec(0, bold = TRUE, color = "white", background = "#2c3e50") # Header styling
```

::: {.callout-note title="Outlier Detection"}
**1. Outlier Detection:** The Bonferroni test reports no Studentized residuals with a $p\text{-value} < 0.05$.
This confirms there are no statistically significant outliers in the dataset.

**2. Influence & Robustness:** Diagnostic analysis based on **Cook's Distance** and **Leverage** further
confirms that no specific observations exert excessive influence on the model's estimates.

**Conclusion:** The model is **robust**. Sensitivity checks suggest that excluding the most extreme
observations would yield very similar coefficients and conclusions. Therefore, we can be confident that our
results are driven by general trends, not by a few exceptional student profiles.
:::

# Interpretation and inference

## Model equation

$$
\begin{aligned}
Y_i =\;& \beta_0 \\
& + \beta_1\,\text{study\_hrs}_i 
  + \beta_2\,\text{sleep\_hrs}_i 
  + \beta_3\,\text{attend\_pct}_i \\
& + \gamma_2\,\mathbb{1}(\text{school\_type}_i = 2) 
  + \sum_{l=2}^{6} \delta_l\,\mathbb{1}(\text{parent\_educ}_i = l) \\
& + \sum_{m=2}^{3} \theta_m\,\mathbb{1}(\text{sleep\_qual}_i = m) 
  + \eta_2\,\mathbb{1}(\text{web\_access}_i = 2) \\
& + \sum_{t=2}^{4} \rho_t\,\mathbb{1}(\text{trav\_time}_i = t) 
  + \kappa_2\,\mathbb{1}(\text{extra\_act}_i = 2) 
  + \sum_{s=2}^{6} \lambda_s\,\mathbb{1}(\text{study\_method}_i = s) \\
& + \sum_{m=2}^{3} \phi_m \bigl(\text{sleep\_hrs}_i \times \mathbb{1}(\text{sleep\_qual}_i = m)\bigr) \\
& + \sum_{m=2}^{3} \psi_m \bigl(\text{study\_hrs}_i \times \mathbb{1}(\text{sleep\_qual}_i = m)\bigr) \\
& + \varepsilon_i
\end{aligned}
$$

$$
\text{where } \varepsilon_i \overset{i.i.d.}{\sim} \mathcal{N}(0, \sigma^2)
$$

$$
\text{Subject to reference constraints: } \gamma_1 = \delta_1 = \theta_1 = \eta_1 = \rho_1 = \kappa_1 = \lambda_1 = \phi_1 = \psi_1 = 0
$$

Yi means the grade of student i, school_typei = 2 means that the school type of student i is private.

## Coefficient table of our model

```{r}

mod_coef <- model_parameters(mod_final, ci_method = "residual", digits = 1)

mod_coef |>
  kable(
    caption = "<strong style='font-size: 20px;'>Coefficient Table</strong>",
    align = "c"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE,
    position = "center"
  ) |>
  scroll_box(height = "400px")
```

## Interpretation of Key Coefficients

**Study_hrs**

Holding all other variables constant, an additional hour of study is associated with an average increase of
about **1.3 points** in the grade for individuals with poor quality of sleep (Poor being the reference of
sleep_qual). This effect represents the baseline marginal effect of study hours, which applies across
categories ***before accounting for interaction terms***.

**Sleep_hrs**

At the referential level of sleep quality (Poor), one additional hour of sleep is associated with an average
increase of approximately **4.8 points** in the grade, all else being equal. This effect represents the
baseline marginal effect of sleep hours, which applies across categories ***before accounting for interaction
terms***.

**trav_time**

Compared to students with a short commute (less than 15 minutes), those traveling between 30 and 60 minutes
score on average about **1.8 points** lower, while students with commutes longer than 60 minutes score
approximately **3.6 points** lower, all else equal. Travel times between 15 and 30 minutes are not associated
with a statistically significant difference relative to the baseline.

Since the model does not include interaction terms involving travel time, ***these estimates can be
interpreted as common effects that apply uniformly across individuals***, regardless of their other
characteristics.

**sleep_qual**

Relative to the baseline category of poor sleep quality, individuals with average sleep quality score about
**12.5 points higher**, while those with good sleep quality score approximately **18.8 points higher**,
holding all other covariates constant.

These coefficients capture baseline differences in performance across sleep-quality categories, independently
of sleep duration.

**sleep_hrs:sleep_qual**

The baseline interpretation of sleep hours corresponds to the reference category of sleep quality. However,
the inclusion of interaction terms between sleep duration and sleep quality indicates that the marginal effect
of an additional hour of sleep is not constant across individuals. (i.e. in a plot of sleep hours against the
grade, the slope of the relationship differs across sleep-quality categories.)

Specifically, the interaction terms show that the positive association between sleep hours and the grade
decreases as sleep quality improves: One additional hour of sleep for people with poor quality of sleep is
associated with an increase of **4.8 points** in the grade (referential category) while one additional hour of
sleep for people with good quality of sleep is associated with an increase of **0.5 point** in the grade
(baseline effect + interaction effect: +4.8 - 4.3 = + 0.5).

While additional sleep is strongly associated with higher performance among individuals with poor sleep
quality, this effect becomes smaller for those with average sleep quality and is substantially attenuated for
individuals reporting good sleep quality.

As a result, ***the overall effect of sleep duration must be interpreted conditionally on sleep quality,
rather than as a single common effect***. These findings suggest diminishing returns to additional sleep when
sleep quality is already high, whereas increasing sleep duration appears most beneficial for individuals
experiencing poor sleep quality.

## Scenario based interpretation

We created two fictional individuals in order to evaluate the predictive abilities of our model. We also
tested for the first individual of the data

```{r}
 new_students <- data.frame(
  id = c(5001, 5002),
  age = c(15, 18),
  sexe = c("Male","Other"),
  study_hrs = c(7, 14),
  sleep_hrs = c(7, 9),
  attend_pct = c(66, 93),
  school_type = c("Public", "Private"),
  parent_educ = c("High School", "Post Graduate 2"),
  sleep_qual = c("Average", "Good"),
  web_access = c("Yes","Yes"),
  trav_time = c(">60 Min", "<15 Min"),
  extra_act = c("No", "Yes"),
  study_method = c("Notes", "Mixed")
)


preds <- predict(mod_final, newdata = new_students  , interval = "confidence")
  


preds |>
  kable(
    caption = "<strong style='font-size: 20px;'>Prediction on new students</strong>",
    align = "c"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE,
    position = "center"
  )

student_1 <- data_clean %>%
  filter(id == 1)

#student_1

pred_student <- predict(mod_final, student_1, interval = "confidence")
pred_student|>
  kable(
    caption = "<strong style='font-size: 20px;'>Prediction on first studend of the original data_set</strong>",
    align = "c"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE,
    position = "center"
  )
```

::: {.callout-tip title="Model Validation: Coherence & Reliability"}
The model demonstrates logical consistency by predicting significantly higher grades for the 'good student'
profile compared to the 'bad student'. Additionally, the fact that the actual grade for Student ID 1 falls
within our predicted confidence interval further validates the model's reliability
:::

## Analysis of Practical Magnitude

While many variables are statistically significant, their real-world impact varies drastically.

Low-impact predictors like study hours (+1.3 points/hr) and school type (+2.0 points) offer only modest gains
relative to the effort or cost involved.

In contrast, Attendance is a transformative driver: despite a small coefficient (β=0.4), the cumulative
difference between low and high attendance can sway grades by 20 points, far outweighing school choice.

Furthermore, daily habits trump background: excellent sleep quality boosts performance by \~18 points, whereas
high parental education adds only \~1.2 points.

Finally, strategy matters: 'Mixed' study methods add value, while 'Group Study' negatively impacts grades
(-1.3 points), suggesting it often serves as a distraction rather than an aid.

## Summarize overall model fit

```{r}
# 1. Extract global statistics
model_fit <- glance(mod_final)

# 2. Select and Rename the 4 Key Metrics
fit_table <- model_fit |>
  select(
    `R-Squared` = r.squared,
    `Adj. R-Squared` = adj.r.squared,
    `Residual Std. Error` = sigma,
    `F-Statistic` = statistic
  )

# 3. Create the Visual Table
fit_table |>
  kable(
    caption = "Global Model Performance Metrics",
    digits = 3,  # Round to 3 decimal places
    align = "cccc"
  ) |>
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"), 
    full_width = FALSE,
    position = "center"
  ) |>
  column_spec(1:4, width = "3cm")
```

# To end in beauty

## Calibrating plot

```{r}
predicted_values <- predict(mod_final, newdata = data_test)

plot_data <- data.frame(
  pred_val = predicted_values,
  obs_val = data_test$y
)

ggplot(plot_data, aes(x = pred_val, y = obs_val)) +
  # Add the points
  geom_point(alpha = 0.5, color = "steelblue", size = 2) +
  
  # Add the perfect prediction line (y = x)
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red", size = 1) +
  
  # Add labels and theme
  labs(
    title = "Calibration Check: Predicted vs. Observed",
    subtitle = "Points on the red dashed line represent perfect predictions",
    x = "Predicted Grade",
    y = "Observed Grade"
  ) +
  theme_minimal() +
  coord_fixed()

```
